\ttitle

### 2015-08-09 Chapters/Introduction.tex
Major changes as below
- Thesis summary completed
- Many changes to thesis.cls and main.tex to format abstract page and university url, etc
- Update Intro-body comp - measuring body comp
- Zot.bin backed up and updated
- Conclusion chapter title renamed to Discussion
- It is 01:40 am - Have listened to Dr. Dre, Snoop Dogg and Eminem all day
- 2 shots of Jura and I am knackered yet sleepless after 2 strong coffees

Patients who develop severe postoperative complications after pancreaticoduodenectomy for cancer had significantly shortened survival in a study involving 428 patients (16.5 vs. 12.4 months, p=0.002) and this was independent of other recognised risk factors such as tumour grade and lymph node status. \parencite{kamphues_postoperative_2011} Similar finding were reported by Raut and co-workers in their study of 360 patients who underwent pancreaticoduodenectomy for pancreatic ductal adenocarcinoma \parencite{raut_impact_2007} and by Kang and co-workers in their report on 103 patients undergoing R0 resections for cancer of the pancreatic head. \parencite{kang_detrimental_2009}

@@

\section{Body Composition, Obesity and Outcomes}
\todo{Introduction - body composition}

The Body Mass Index (BMI) is a measure of a person's weight relative to their height and corresponds well with their body composition, especially with the proportion that is body fat. It is the most widely accepted measure of body composition, especially of body fat in spite of criticisms that it may not apply to all populations. BMI was first devised by Adolphe Quetelet and was formerly known as the Quetelet index.\parencite{eknoyan_adolphe_2008} Quetelet was a Belgian sociologist and mathematician who played a major role in the application of statistical methods to the social sciences. In his attempts to describe the characteristics of 'normal man', Quetelet found that weight increased nearly as the square of the height except at times of growth spurts in the new born and during puberty. \parencite{quetelet_treatise_2013} The calculation of body mass index is simple and as follows:

@@

\begin{table}[h]
	\centering
	\caption{The WHO classification of Body Mass Index}
	\label{table:bmi_who}
	\begin{tabular}{ |l|l|}
		\hline
		BMI         & Category                  \\ \hline
		Below 18.5  & Underweight               \\
		18.5 – 24.9 & Normal range              \\
		25.0 – 29.9 & Overweight or Pre-obesity \\
		30.0 - 34.9 & Obesity class 1             \\
		35.0 - 39.9 & Obesity class 2             \\
		Above 40    & Obesity class 3             \\ \hline
	\end{tabular}
\end{table}

@@

Several studies have reported increased incidence of postoperative complications after cardiac, transplant, urologic and gynaecological surgery. \parencite{prem_operative_1965, holley_obesity_1990, fasol_influence_1992, gruberg_impact_2002, lee_impact_2004} Visceral obesity rather than body mass index has been reported to be associated with greater incidence of wound complications, overall complication rate and prolonged hospital stay in patients undergoing laparascopic sigmoid colectomy for cancer. \parencite{tsujinaka_visceral_2008}

@@

In a study of 356 consecutive patients undergoing pancreaticoduodenectomy at the Memorial Sloan-Kettering Cancer Center between 2000 and 2005, House and co-workers used preoperative cross-sectional imaging of the abdomen (computer tomography and/or magnetic resonance imaging) to measure a variety of indices of body composition including abdominal wall fat thickness, hip girdle fat thickness, retro-renal visceral fat and pancreatic duct diameter. \parencite{house_preoperative_2008} The incidence of wound infection was greater in obese patients with the risk increasing linearly with BMI. But, BMI was not associated with any other complications. They also reported that the thickness of retro-renal fat, a surrogate for visceral adipose tissue, was a significant risk factor for pancreatic fistula as well as wound infection. However, there was no correlation between complications and abdominal wall thickness or hip girdle fat. The importance of visceral fat in the incidence and severity of postoperative complications is further emphasised by the finding that excess visceral fat is associated with the the occurence of pancreatic fistula after radical total gastrectomy for gastric cancer. Benns and co-workers reported greater blood loss and increased incidence and severity of postoperative complications in obese pateints undergoing surgery compared to non-obese pateints. However, there was no difference in operative time or length of stay in hospital. \parencite{benns_impact_2009}

@@

>>>>

Patients who develop severe postoperative complications after pancreaticoduodenectomy for cancer had significantly shortened survival in a study involving 428 patients (16.5 vs. 12.4 months, p=0.002) and this was independent of other recognised risk factors such as tumour grade and lymph node status. \parencite{kamphues_postoperative_2012} Similar finding were reported by Raut and co-workers in their study of 360 patients who underwent pancreaticoduodenectomy for pancreatic ductal adenocarcinoma \parencite{raut_impact_2007} and by Kang and co-workers in their report on 103 patients undergoing R0 resections for cancer of the pancreatic head. \parencite{kang_detrimental_2009}

@@

\section{Body Composition, Obesity and Outcomes}
\todo{Introduction - body composition}
\subsection{Body Mass Index}

The Body Mass Index (BMI) is a measure of a person's weight relative to their height and corresponds well with their body composition, especially with the proportion that is body fat. It is the most widely accepted measure of body composition, especially of body fat in spite of criticisms that it may not apply to all populations. BMI was first devised by Adolphe Quetelet and was formerly known as the Quetelet index.\parencite{eknoyan_adolphe_2008} Quetelet was a Belgian sociologist and mathematician who played a major role in the application of statistical methods to the social sciences. In his attempts to describe the characteristics of 'normal man', Quetelet found that weight increased nearly as the square of the height except at times of growth spurts in the new born and during puberty. \parencite{quetelet_treatise_1842} The calculation of body mass index is simple and as follows:

@@

\begin{table}[h]
	\centering
	\caption{The WHO classification of Body Mass Index}
	\label{table:bmi_who}
	\renewcommand{\arraystretch}{1.5} %Increases space between rows
	\begin{tabular}{|l l|}
		\hline
		BMI         & Category                  \\ \hline
		Below 18.5  & Underweight               \\
		18.5 – 24.9 & Normal range              \\
		25.0 – 29.9 & Overweight or Pre-obesity \\
		30.0 - 34.9 & Obesity class 1             \\
		35.0 - 39.9 & Obesity class 2             \\
		Above 40    & Obesity class 3             \\ \hline
	\end{tabular}
\end{table}

@@

Several studies have reported increased incidence of postoperative complications in obese patients after cardiac, transplant, urologic and gynaecological surgery. \parencite{prem_operative_1965, holley_obesity_1990, fasol_influence_1992, gruberg_impact_2002, lee_impact_2004} Visceral obesity rather than body mass index has been reported to be associated with greater incidence of wound complications, overall complication rate and prolonged hospital stay in patients undergoing laparascopic sigmoid colectomy for cancer. \parencite{tsujinaka_visceral_2008}

@@

In a study of 356 consecutive patients undergoing pancreaticoduodenectomy at the Memorial Sloan-Kettering Cancer Center between 2000 and 2005, House and co-workers used preoperative cross-sectional imaging of the abdomen (computer tomography and/or magnetic resonance imaging) to measure a variety of indices of body composition including abdominal wall fat thickness, hip girdle fat thickness, retro-renal visceral fat and pancreatic duct diameter. \parencite{house_preoperative_2008} The incidence of wound infection was greater in obese patients with the risk increasing linearly with BMI. But, BMI was not associated with any other complications. They also reported that the thickness of retro-renal fat, a surrogate for visceral adipose tissue, was a significant risk factor for pancreatic fistula as well as wound infection. However, there was no correlation between complications and abdominal wall thickness or hip girdle fat.

The importance of visceral fat in the incidence and severity of postoperative complications is further emphasised by the finding that excess visceral fat is associated with the the occurence of pancreatic fistula after radical total gastrectomy for gastric cancer. Benns and co-workers reported greater blood loss and increased incidence and severity of postoperative complications in obese pateints undergoing surgery compared to non-obese pateints. However, there was no difference in operative time or length of stay in hospital. \parencite{benns_impact_2009}

@@

While sarcopenia or loss of skeletal muscle mass is associated with poor outcome in patients with pancreatic cancer, the association between adipose tissue (especially, visceral fat) and long-term outcomes is less clear. Sarcopenic-obesity where concurrent sarcopenia and high fat mass are present is associated with significantly worse survival when compared to sarcopenia or high adiposity on their own. \parencite{tan_sarcopenia_2009} It is clear that body mass index on its own is a poor predictor of outcome and body composition changes over time play a more important role in determining cancer survival. While obesity appears to have a protective role in a wide range of chronic disease states\parencite{kalantar-zadeh_risk_2007, amundson_obesity_2010, myers_obesity_2011} sarcopenic-obesity has a negative impact on survival. \parencite{tan_sarcopenia_2009}

\subsection{Measuring Body Composition}
While it is possible to estimate the amount of most tissue types, the ones that are of clinical interest and are related to surgical and cancer outcomes are adipose tissue and skeletal muscle. Numerous methods have been described for estimation of both absolute and proportional amount of these tissues with varying degrees of complexity and accuracy. Some of the simpler methods include measurement of subcutaneous fat using calipers or bioelectrical impedance, estimation of skeletal muscle from mid-arm circumference or creatinine height ratio. Such methods are valuable to public health and epidemiological studies of malnutrition.

The gold-standard method for assessment of body composition is Dual Energy X-ray Absorptiometry (DEXA).\parencite{haarbo_validation_1991} While primarily designed for assessment of bone mineral density, DEXA scans have been adapted to measure fat, muscle and water content of subjects accurately and are used widely both in research and clinical settings. Other techniques such as whole body water or air displacement plethysmography are relatively cumbersome for routine clinical use.

More recently, cross-sectional imaging including computed tomography (CT) and magnetic resonance imaging (MRI) have become widely available. Nearly all patients with gastrointestinal cancer undergo either CT or MRI or both before commencing treatment. Accurate measurements of subcutaneous adipose tissue, visceral adipose tissue and skeletal muscle can be made relatively easy either from the entire series of images or estimates can be made from measurements made at certain predetermined anatomical levels. Both specialised software or readily available generic image analysis tools have been used to make these measurements.

The proportion of the total cross-sectional area that is made of adipose tissue or skeletal muscle at the level of the 3$^{rd}$  or the 4$^{th}$ lumbar vertebra is commonly used as a surrogate for total body measurements. Thickness of retro-renal visceral fat at the level of the renal vein has also been reported to be usable as a surrogate for total visceral adiposity.

### 2015-08-09 main.tex
Introduction - Obesity and Survival
Major changes as below
- Thesis summary completed
- Many changes to thesis.cls and main.tex to format abstract page and university url, etc
- Update Intro-body comp - measuring body comp
- Zot.bin backed up and updated
- Conclusion chapter title renamed to Discussion
- It is 01:40 am - Have listened to Dr. Dre, Snoop Dogg and Eminem all day
- 2 shots of Jura and I am knackered yet sleepless after 2 strong coffees

\abstract{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics

To be finalised...
}
\clearpage

@@

% Include the chapters of the thesis as separate files from the Chapters folder
\input{Chapters/Introduction} 		%1
\input{Chapters/CPET_Outcomes} 	%2
\input{Chapters/CPET_Jaundice} 	%3
\input{Chapters/BodyComp}			%4
\input{Chapters/CPET_Survival}		%5
\input{Chapters/CRP_Complications}	%6
\input{Chapters/Discussion}		%7

@@

\input{Appendices/AppendixExcelMacros}
\input{Appendices/AppendixCPETRawData}
%\input{Appendices/AppendixC}

>>>>

\abstract{\addtocontents{toc}{\vspace{1em}} % Add a gap in the Contents, for aesthetics
Pancreaticoduodenectomy with or without adjuvant chemotherapy remains the only modality of possible cure in patients with cancer involving the head of the pancreas and the periampullary region. While mortality rates after pancreaticoduodenectomy have improved considerably over the course of the last century, morbidity remains high. Patient selection is of paramount importance in ensuring that major surgery is offered to individuals who will most benefit from a pancreaticoduodenectomy. Moreover, identifying preoperative risk factors provides potential targets for prehabilitation and optimisation of the patient's physiology before undertaking surgery. In addition to this, early identification of patients who are likely to develop postoperative complications allows for better allocation of critical care resources and more targeted treatment of the high risk patients.

Cardiopulmonary exercise testing is becoming an increasingly popular tool in the preoperative risk assessment of the surgical patient. However, very little work has been done to investigate the role of cardiopulmonary exercise test in patients scheduled to undergo pancreaticoduodenectomy. The impact of jaundice, sytemic inflammation and other preoperative clinicopathological characteristics of the patient on their cardiopulmonary exercise physiology has not been studied in detail before in this cohort of patients. Moreover, the impact of these factors on long-term survival is also unknown.

The overall aim of the thesis was to evaluate the relationship between cardiopulmonary exercise testing, perioperative clinico-pathological characteristics and postoperative outcomes as well as long-term survival in patients undergoing pancreaticoduodenectomy.

\textbf{Chapter \ref{ch_intro}} reviews the existing literature on preoperative cardiopulmonary exercise testing, the impact of obstructive jaundice, perioperative systemic inflammation and the importance of body composition in determining outcomes in patients undergoing major surgery with particular reference to pancreatic surgery.

\textbf{Chapter \ref{ch_cpet_outcomes}} reports on the role of cardiopulmonary exercise testing in predicting postoperative complications. The results demonstrate that patients with anaerobic threshold less than 10 mls/kg/min were more likely to develop a postoperative pancreatic fistula, stay longer in hospital and less likely to receive adjuvant therapy after pancreaticoduodenectomy. These results emphasise the importance of excellent aerobic fitness to improve postoperative outcomes and provide a potential target for future prehabilitation programmes.

\textbf{Chapter \ref{ch_cpet_jaundice}} evaluates the relationship between cardiopulmonary exercise physiology and other clinicopathological characteristics of the patient. In a detailed analysis of cardiopulmonary exercise test parameters in jaundiced versus non-jaundiced patients, the results in this chapter demonstrate that obstructive jaundice does not impair cardiopulmonary exercise physiology. This further supports emerging evidence in contemporary literature that jaundiced patients can proceed directly to surgery without underoing preoperative biliary drainage. The results of this study also show an interesting inverse relationship between body mass index and anaerobic threshold which is analysed in more detail in Chapter 3.

\textbf{Chapter \ref{ch_bodycomp}} examines the relationship between preoperative cardiopulmonary exercise physiology and body composition in depth. All parameters measured at cardiopulmonary exercise test are compared against body composition and body mass index. The results of this chapter report that the current method of reporting $\dot{V}_{O_2}$, both at peak exercise and anaerobic threshold, is biased against obese subjects and advises caution in the interpretation of cardiopulmonary exercise test results in patients with a high BMI. This is particularly important as current evidence in literature suggests that postoperative outcomes in obese subjects are comparable to non-obese subjects while cardiopulmonary exercise test results are also abnormally low in this very same cohort of patients.

\textbf{Chapter \ref{ch_survival}} analyses the relationship between preoperative comorbidity including cardiopulmonary exercise test and long-term survival in patients with pancreatic ductal adenocarcinoma undergoing pancreaticoduodenectomy. The results of this chapter suggest that cardiopulmonary exercise testing parameters are not useful in predicting long-term survival. However, a composite score of preoperative physiology as measured by the POSSUM Physiology score was associated with survival.

\textbf{Chapter \ref{ch_crp_comp}} evaluates the role of the early postoperative systemic inflammatory response in predicting complications after pancreaticoduodenectomy and aims to establish clinically relevant thresholds for C-Reactive Protein for the prediction of complications. The results of this chapter demonstrate that CRP levels as early as the second postoperative day are associated with complications. While post-operative CRP was useful in the prediction of infectious complications, this was only the case in patients who did not develop a post-operative pancreatic fistula. The predictive ability of inflammatory markers for infectious complications was blunted in patients with a pancreatic fistula. In these patients, outcome was largely determined by the severity of the pancreatic fistula.

\textbf{Chapter \ref{ch_discussion}} summarises the findings of this thesis, their place in current literature and future directions. The contents of this thesis aim to add to the current body of knowledge regarding the complex pathophysiological alterations that occur in patients with lesions of the head of the pancreas with the associated obstructive jaundice, systemic inflammation and their impact on cardiopulmonary exercise physiology, postoperative complications and long-term outcomes. The results establish the value of cardiopulmonary exercise test as an objective measure of aerobic fitness while refuting a commonly held belief that obstructive jaundice significantly affected cardiovascular function. The current work paves the way for further study into evaluating the role of cardiopulmonary exercise test in prehabilitation regimens as well as the incorporation of preoperative $\dot{V}_{O_2}$AT and postoperative CRP into existing models for predicting complications after pancreaticoduodenectomy. These results will help improve patient selection and preparation for pancreaticoduodenectomy  and ultimately the short and long-term outcomes after this morbid yet necessary operation.

}
\clearpage

@@

% Include the chapters of the thesis as separate files from the Chapters folder
\input{Chapters/Introduction} 		%1
%\input{Chapters/CPET_Outcomes} 	%2
%\input{Chapters/CPET_Jaundice} 	%3
%\input{Chapters/BodyComp}			%4
%\input{Chapters/CPET_Survival}		%5
%\input{Chapters/CRP_Complications}	%6
%\input{Chapters/Discussion}		%7

@@

%\input{Appendices/AppendixExcelMacros}
%\input{Appendices/AppendixCPETRawData}
%\input{Appendices/AppendixC}
